`SimSurvey`: an `R` package for comparing the design and analysis of fisheries surveys by simulating spatially-correlated fish stocks

### 2020-02-12 analysis/paper/sim_survey_paper.Rmd
Wrap up draft interpretation section and add a sentence to the abstract

Fish populations often show complex spatial and temporal dynamics, creating challenges in designing and implementing effective surveys. Inappropriate sampling designs can potentially lead to both under-sampling (reducing precision and increasing the risk of bias) and over-sampling (through the extensive and potentially expensive sampling of correlated metrics). For assessments based on fisheries-independent surveys, the ability to estimate population parameters is affected by multiple levels of sampling, such as the number of sampling stations as well as the sub-sampling of fish captured to measure biological characteristics (e.g. lengths or ages). Population estimates are also affected by the pathway taken to analyze such data. Though simulations are a useful tool for exploring the efficacy of specific sampling strategies and statistical methods, there are a limited the number of tools that facilitate the simulation testing of a range of sampling and analytical pathways for fisheries-independent survey data. Here we introduce the `R` package **`SimSurvey`**, which has been designed to simplify the process of simulating surveys of age-structured and spatially-distributed fish populations. The package allows the user to simulate age-structured populations that vary in space and time and explore the efficacy of a range of built-in or user-defined sampling protocols to reproduce the population parameters of the known population. **`SimSurvey`** also includes a function for estimating the stratified mean and variance of the population from the simulated survey data. **`SimSurvey`** can serve as a convenient, accessible and flexible platform for simulating a wide range of sampling strategies for fish stocks that show complex structuring. Various statistical approaches can then be applied to the results to test the efficacy of different analytical approaches.

@@

This simulation framework has similarities to those presented by Schnute and Haigh [-@schnute2003] and Puerta et al. [-@puerta2019], however, efforts were focused on developing a series of general and accessible functions to simplify the process of testing multiple sampling scenarios and analytical pathways. The steps taken to simulate surveys of spatial, age-structure populations are outlined below. Under the **[Methods]** section, we first outline the equations underlying the package (**[Model structure]** section) and then move into a vignette of the package (**[Core functions]** section). While the core of this paper focuses on how to use this package, it is important to note that the defaults of the package are based on a case study. Output from default function calls are therefore relevant to the case study and some of these results are highlighted in the **[Results]** and **[Discussion]** section; a more detailed description and discussion of these results can be found in **[S1 Appendix][S1 Appendix: Case study]**.

@@

Results shown in the **[Core functions]** section demonstrates one use case of the **`SimSurvey`** package that is focused on evaluating the efficacy of increasing the sampling effort of a stratified random survey of a cod population (see **[S1 Appendix][S1 Appendix: Case study]** for details). These results largely align with expectations from sampling theory that 1) design-based estimators for stratified random surveys are unbiased, and 2) precision is increased by increasing the number of primary sampling units [@Cochran1977]. Specifically, estimates of total abundance and abundance at length are centered around true values and their probability envelopes tighten as set density increases (Figures \@ref(fig:fan1), \@ref(fig:fan2)). Case study results also echo the growing body of literature which concludes that extra sub-sampling is an ineffective means of improving estimates relative to sampling more locations [@Pennington1994; @Pennington2002; @Stewart2014; @Bogstad1995; @Coggins2013; @Zhang2013]. This is exemplified by the relatively large drops in RMSE when set density is increased compared to when sub-sampling effort is increased (Figures \@ref(fig:rank), \@ref(fig:rmse-surface)). Moreover, it appears that it is more advantageous to measure fewer total fish at more locations than measuring many fish at fewer locations (Figure \@ref(fig:rank)). Again, this result was expected because the size-structured spatial clustering imposed by the simulation causes set-to-set variation to be greater than within-set variation in characteristics such as length and age. This structure was imposed because fish caught together tend to have similar characteristics and it is well known that this reduces the effective sample size [@Pennington1994]. Increases to the number of sub-samples, however, should not lead to poorer population estimates like those observed for abundance at age (Figure \@ref(fig:fan3), \@ref(fig:rmse-surface)). Bias, in particular, appears to be a problem with stratified estimates of abundance at age under certain scenarios (Figure \@ref(fig:fan3)); in contrast, there appears to be little bias in the stratified estimates of abundance at length (Figure \@ref(fig:fan2)). This result was unexpected as the design-based analysis was presumed to be unbiased. The contrast between the length and age based analysis indicates that the problem lies with the intervening age-length-key.

Age-length keys have long been used in conjunction with length frequency tables to estimate the age distribution of fish populations over large scales [@fridriksson1934]. The default option applied in the package is to construct and apply the key at the division level, as this is the standard protocol for the actual survey the case study is based on. There is, however, a potential cost to the spatial scale of the key. Namely, it is unlikely that one age-length key is representative for the whole region because the probability of being a specific age given length varies in space [@Berg2012]. Because different age groups sometimes occur in different places, the translation of lengths to ages may be biased by the samples used to generate the age-length key. In short, multi-stage sampling of a clustered population violates the assumption that the age-length key is generated from a simple random sample [@Aanes2015]. Bias introduced by this assumption can be resolved by explicitly accounting for hierarchical sampling using designed-based estimators [@Aanes2015]. For the case study, the estimator will need to account for cluster sampling in a stratified random survey. Following recommendations in Aanes and Vølstad [@Aanes2015], we used **`SimSurvey`** to 1) conduct a surveys with concurrent length and age sampling at every set, 2) construct and apply age-length keys on a set-by-set basis, and 3) weight the resultant age frequencies at each set using stratified random estimators. This test was accomplished using this code:

@@

```{r alt-tests, out.height = "600px", fig.cap = "Key results from an alternative design-based sampling and analysis procedure that accounts for multi-stage cluster sampling of ages. Specifically, a) is a fan chart of stratified estimates of abundance at age from year seven of the simulation from a survey with a set density ($D_{\\mathrm{sets}}$) of 0.002 sets / km^2^, maximum number of length samples ($M_{\\mathrm{lengths}}$) of 100, and maximum number of age samples per 3 cm length group per set ($M_{\\mathrm{ages}}$) of 3 (the thick black line indicates the true trend in the total population available to the survey and the yellow to purple color gradient represents a range of probability envelopes from 10% to 90%); and b) is a surface plot of RMSE from an array of surveys with different sub-sampling effort, where the x-axes represent the maximum number of lengths to measure per set ($M_{\\mathrm{lengths}}$), and y-axes represent the maximum number of ages to collect per 3 cm length group per set ($M_{\\mathrm{ages}}$)."}

@@

These estimators require concurrent length and age sampling at every primary sampling unit

This bias is perhaps compounded by the length-stratified sampling of ages such that the ages sampled may be skewed towards sets with the most catch, especially under scenarios where length sampling effort is high. If this is the case, then the representativeness of the age-length key to the whole population could be diminished by excessive length sampling.

Age-length-key...large scale typical...danger, assumptions...segue to solution...

>>>>

Fish populations often show complex spatial and temporal dynamics, creating challenges in designing and implementing effective surveys. Inappropriate sampling designs can potentially lead to both under-sampling (reducing precision) and over-sampling (through the extensive and potentially expensive sampling of correlated metrics). For assessments based on fisheries-independent surveys, the ability to estimate population parameters is affected by multiple levels of sampling, such as the number of sampling stations as well as the sub-sampling of fish captured to measure biological characteristics (e.g. lengths or ages). Population estimates are also affected by the pathway taken to analyze such data. Though simulations are a useful tool for exploring the efficacy of specific sampling strategies and statistical methods, there are a limited the number of tools that facilitate the simulation testing of a range of sampling and analytical pathways for fisheries-independent survey data. Here we introduce the `R` package **`SimSurvey`**, which has been designed to simplify the process of simulating surveys of age-structured and spatially-distributed fish populations. The package allows the user to simulate age-structured populations that vary in space and time and explore the efficacy of a range of built-in or user-defined sampling protocols to reproduce the population parameters of the known population. **`SimSurvey`** also includes a function for estimating the stratified mean and variance of the population from the simulated survey data. We demonstrate the use of this package using a case study and show that the it can reveal unexpected sources of bias and be used to explore design-based solutions to such problems. In summary, **`SimSurvey`** can serve as a convenient, accessible and flexible platform for simulating a wide range of sampling strategies for fish stocks that show complex structuring. Various statistical approaches can then be applied to the results to test the efficacy of different analytical approaches.

@@

This simulation framework has similarities to those presented by Schnute and Haigh [-@schnute2003] and Puerta et al. [-@puerta2019], however, efforts were focused on developing a series of general and accessible functions to simplify the process of testing multiple sampling scenarios and analytical pathways. The steps taken to simulate surveys of spatial, age-structure populations are outlined below. Under the **[Methods]** section, we first outline the equations underlying the package (**[Model structure]** section) and then move into a vignette of the package (**[Core functions]** section). While the core of this paper focuses on how to use this package, it is important to note that the defaults of the package are based on a case study. Output from default function calls are therefore relevant to the case study and some of these results are highlighted in the **[Interpretation]** section; a more detailed description and discussion of these results can be found in **[S1 Appendix][S1 Appendix: Case study]**.

@@

Results shown in the **[Core functions]** section demonstrates one use case of the **`SimSurvey`** package that is focused on evaluating the efficacy of increasing the sampling effort of a stratified random survey of a cod population (see **[S1 Appendix][S1 Appendix: Case study]** for details). These results largely align with expectations from sampling theory that 1) design-based estimators for stratified random surveys are unbiased, and 2) precision is increased by increasing the number of primary sampling units [@Cochran1977]. Specifically, estimates of total abundance and abundance at length are centered around true values and their probability envelopes tighten as set density increases (Figures \@ref(fig:fan1), \@ref(fig:fan2)). Case study results also echo the growing body of literature which concludes that extra sub-sampling is an ineffective means of improving estimates relative to sampling more locations [@Pennington1994; @Pennington2002; @Stewart2014; @Bogstad1995; @Coggins2013; @Zhang2013]. This is exemplified by the relatively large drops in RMSE when set density is increased compared to when sub-sampling effort is increased (Figures \@ref(fig:rank), \@ref(fig:rmse-surface)). Moreover, it appears that it is more advantageous to measure fewer total fish at more locations than measuring many fish at fewer locations (Figure \@ref(fig:rank)). Again, this result was expected because the size-structured spatial clustering imposed by the simulation causes set-to-set variation to be greater than within-set variation in characteristics such as length and age. This structure was imposed because fish caught together tend to have similar characteristics and it is well known that this reduces the effective sample size [@Pennington1994]. Increases sub-sampling effort, however, should not lead to poorer population estimates like those observed for abundance at age (Figure \@ref(fig:fan3), \@ref(fig:rmse-surface)). Bias, in particular, appears to be a problem with stratified estimates of abundance at age under certain scenarios (Figure \@ref(fig:fan3)); in contrast, there appears to be little bias in the stratified estimates of abundance at length (Figure \@ref(fig:fan2)). This result was unexpected as the analysis was presumed to be unbiased. The contrast between the length and age based analysis indicates that the problem lies with the intervening age-length-key.

Age-length keys have long been used in conjunction with length frequency tables to estimate the age distribution of fish populations over large scales [@fridriksson1934]. By default, `sim_survey` spreads length-stratified age sampling across the division and `run_strat` constructs and applies age-length keys at the division scale. This default was chosen because it is standard protocol for the actual survey the case study is based on. There is, however, a potential cost to the spatial scale of the key. Namely, it is unlikely that one age-length key is representative for the whole region because the probability of being a specific age given length varies in space [@Berg2012]. Because different age groups sometimes occur in different places, the translation of lengths to ages may be biased by the samples used to generate the age-length key. In short, multi-stage sampling of a clustered population violates the assumption that the age-length key is generated from a simple random sample [@Aanes2015]. Bias introduced by this assumption can be resolved by explicitly accounting for the hierarchical sampling by using designed-based estimators [@Aanes2015]. For the case study, a proper design-based estimator will need to account for cluster sampling in a stratified random survey. Following recommendations in Aanes and Vølstad [@Aanes2015], we used **`SimSurvey`** to 1) conduct a surveys with concurrent length and age sampling at every set, 2) construct and apply age-length keys on a set-by-set basis, and 3) weight the resultant age frequencies at each set using stratified random estimators. This test was accomplished using this code:

@@

```{r alt, out.height = "600px", fig.cap = "Key results from an alternative design-based sampling and analysis procedure that accounts for multi-stage cluster sampling of ages. Specifically, a) is a fan chart of stratified estimates of abundance at age from year seven of the simulation from a survey with a set density ($D_{\\mathrm{sets}}$) of 0.002 sets / km^2^, maximum number of length samples ($M_{\\mathrm{lengths}}$) of 100, and maximum number of age samples per 3 cm length group per set ($M_{\\mathrm{ages}}$) of 3 (the thick black line indicates the true trend in the total population available to the survey and the yellow to purple color gradient represents a range of probability envelopes from 10% to 90%); and b) is a surface plot of RMSE from an array of surveys with different sub-sampling effort, where the x-axes represent the maximum number of lengths to measure per set ($M_{\\mathrm{lengths}}$), and y-axes represent the maximum number of ages to collect per 3 cm length group per set ($M_{\\mathrm{ages}}$)."}

@@

Unlike the default approach to collecting and analyzing age samples, results from this test conform to expectations in that the estimator appears to be unbiased and, ignoring the costs associated with unnecessary sampling effort, there appears to be no negative consequences of increasing sub-sampling effort (Figure \@ref(fig:alt)). This is an important result with real-world implications for the survey the case study is based on. This example shows how **`SimSurvey`** can be used to reveal problems and explore solutions to the sampling designs of fisheries-independent surveys.
